Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-04 WSSSPE1_report/WSSSPE1_summary_report.tex
adding sustainability notes in 8.1 resolving a question

He experienced problems with staff retention with respect to private organizations which reward 
those combining research and software expertise (the ``Google Bus''). However,
he is a strong fan of private sector sustainability sources and PPP\note{what does PPP mean?}. He
noted that making a successful business from scientific software alone
is very rare: founders overvalue while customers undervalue. He noted
that to last, an open source project needs a minimal funding requirement
even with a vibrant community -- good will only goes so far if you are
being paid to do something else.

@@

>>>>

He experienced problems with staff retention with respect to private organizations which reward 
those combining research and software expertise (the ``Google Bus''). However,
he is a strong fan of private sector sustainability sources and public-private partnerships. He
noted that making a successful business from scientific software alone
is very rare: founders overvalue while customers undervalue. He noted
that to last, an open source project needs a minimal funding requirement
even with a vibrant community -- good will only goes so far if you are
being paid to do something else.

@@

\item From discussion in Phil Bourne's keynote:
\begin{itemize}
\item Can sustainability be defined as the effort that happens to make the essential things continue?
\item What is sustainability? What is it that we want to sustain? Is what we want to sustain valuable? Who cares if it goes away and how do you measure how much they care?
\end{itemize}

\item from discussion in developing panel

\begin{itemize}
\item what does sustainable mean?
\begin{itemize}
\item from http://oss-watch.ac.uk/resources/ssmm: When choosing software for procurement or development reuse - regardless of the license and development model you will use - you need to consider the future. While a software product may satisfy today's needs, will it satisfy tomorrow's needs? Will the supplier still be around in five years' time? Will the supplier still care for all its customers in five years' time? Will the supplier be responsive to bug reports and feature requests? In other words, is the software sustainable?
\item I think this is a key question we as a community need to agree on (whats our problem statement?) and hope we come out today with some progress. - @VonWelch
\item What is the goal of sustainability? Reproducible science, persistence, quality, or else?
\item How do you measure success? How do you know when you have actually achieved sustainable software?
\item Sustainability should be addressed throughout the life cycle (Daisie Huang)
\end{itemize}

\item What is the relationship between sustainability and provenance?

\item Is usability separate from sustainability or a fundamental part?
\end{itemize}

\item from discussion in policy panel

\begin{itemize}
\item Chris Lenhardt's presentation of Venters paper~\cite{Venters_WSSSPE} included:

\begin{itemize}
\item What does software sustainability mean
It ``is a rather ambiguous concept...''
\item Lack of an accepted definition gets in the way of integrating the concept into software engineering
\item Sustainability as a non-functional requirement
\item Quality of software architectures determines sustainability
\item Propose sustainability is a measure of central quality attributes:
Extensibility,
Interoperability,
Maintainability,
Portability,
Reusability,
Scalability
\item Develop an architecture evaluation framework based on scenarios which help to illuminate how to measure quality/sustainability at the architectural level
\end{itemize}

\item Chris Lenhardt's presentation of Katz paper~\cite{Katz_WSSSPE} included:
\begin{itemize}
\item Clarifying the notion of sustainability
\begin{itemize}
\item Provide same functionality going forward even if environment changes
\item Is functionality and usability clear
\item Will functionality be correct going into the future even if environment changes
\item Is the functionality desired by current and future users
\item Can it incorporate new things as they develop
\end{itemize}
\end{itemize}

\item see slide 9 of \url{http://dx.doi.org/10.6084/m9.figshare.853817} for summary table

\item also see discussion at \url{https://docs.google.com/document/d/1eVfioGNlihXG_1Y8BgdCI6tXZKrybZgz5XuQHjT1oKU/edit?pli=1#bookmark=id.2pxw5yey8jl0}
